---
title: "Quick R Tutorial"
author: "by Sabrina Mai Nielsen"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\  

#Getting started - very quickly

1. Download R from https://www.r-project.org/

2. Open the R program

3. Click 'File' -> 'New Script', and you are ready!

You now have two windows - one is the Editor window (the empty one) - this is where you will be writing code. The other window is the Console window. This is where the code you run will be shown together with the results of calculations. 


```{r, include=FALSE}

library(knitr)    # For knitting document and include_graphics function
library(ggplot2)  # For plotting
library(png)      # For grabbing the dimensions of png files

```

\  
\ 
<center>
```{r echo=FALSE}
img1_path <- "images//Rtuturial_Rscreenshot0.png"
include_graphics(img1_path)
```
</center>

\ 

**General tips:**

* To run the code, type `Ctrl` + `r` on your keyboard - this will execute the code on the line where the marker is placed in the Editor window. You can run all code in a Editor window by typing `Ctrl` + `a`, and then `Ctrl` + `r`.

* When you save your codes in the Editor window, click 'File' -> 'Save as'. Remember to write '.r' after the filename.

* If you want to open an R file, you will need to open it within the program (and not simply by clicking on it), using 'File' -> 'Open script'.

* Save often - R does not save by itself.

\ 

#Get a feeling of R

###Use R as a calculator

Try to type the following in the Editor window:

```{r}
 10+98

 10 + 98

 10+
 98

 sum(10,98)

```

Now, run all the code by typing `Ctrl` + `a`, and then `Ctrl` + `r`.
The results should appear in the Console window.  

As you may notice, R does not care about spaces or line shifts.

\  
\ 

Now your screen should look like this:

<center>
```{r echo=FALSE}
img1_path <- "images//Rtuturial_Rscreenshot1.png"
include_graphics(img1_path)
```
</center>

\ 

Further simple calculations can be written like this:
```{r}
 31-7	

 10*5

 10/5

 sqrt(9)

 3^2
 
 (5+3)*2

```

\ 

In R, it is possible to assign values to letters or word:
```{r}
#Assign value (FYI everything written after '#' will be ignoret and is useful for notes)
 a <- 2
 b <- 5
 a+b

```

\ 

Or make a function:
```{r}
#Funtions
 MyFunction <- function(x){3*x+5}
 MyOtherFunction <- function(z){z^2+50}

 MyFunction(2)
 MyOtherFunction(2)

```

\ 
\ 

###Plots in R

When you want to make plots in R, the function `plot()` is simple and relatively easy to use. The plots will appear in a new window, an R Graph window.
```{r, eval=FALSE}
plot(MyOtherFunction)
```

\  
\ 

Now your screen should look like this:

<center>
```{r echo=FALSE}
img1_path <- "images//Rtuturial_Rscreenshot2.png"
include_graphics(img1_path)
```
</center>

\  

In addition to the functions build into R as default, you can install packages for your R program. First time you use a package, you will need to install it with `install.packages()`. For using the package, you need to load the package everytime you open your R program, with `librarary()`. If you are asked, about CRAN mirrors, you may just pick the one called '0-Cloud [https]'.

```{r, eval=FALSE}
install.packages("ggplot2") #Run only once
library(ggplot2)

ggplot(data.frame(x = c(0, 1)), aes(x = x)) +
        stat_function(fun = MyOtherFunction)
```

```{r, echo=FALSE}
ggplot(data.frame(x = c(0, 1)), aes(x = x)) +
        stat_function(fun = MyOtherFunction)
```

\  

<center>
## [Now you are ready for trying out some graphs in R](R-Graphs.html)
</center>


