---
title: "R Graphs"
author: "by S. M. Nielsen"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}

library(knitr)    # For knitting document and include_graphics function
library(ggplot2)  # For plotting
library(png)      # For grabbing the dimensions of png files

```

\ 

# Line chart with confidence intervals

Let's try to reproduce panel A in this figure published in BMJ Open by <a href="https://www.ncbi.nlm.nih.gov/pubmed/29420232" target="_blank">Roos et al. (2018)</a>, where knee surgery (arthroscopic partial meniscectomy) was compare to sham surgery (skin incision only). The graph shows the mean change from baseline at 3 and 24 months in the primary outcome (KOOS~5~ score, a higher score is better). Further details can be found in the article.

\  
\ 
<center>
```{r echo=FALSE}
img1_path <- "images//FIGURE_Roos_2018.png"
include_graphics(img1_path)
```
</center>

\ 

###Read in the data

To make it simple, all data needed for the figure is typed into R in this format.

```{r FIG_R0, echo=FALSE}
MONTH <- c(0,3,24,0,3,24)
Group <- c("APM","APM","APM","Placebo","Placebo","Placebo")
lsmean <- c(0, 13.9, 21.8, 0, 15.1, 13.6)
lower.CL <- c(0, 3.2, 11.2, 0, 4.2, 3.0)
upper.CL <- c(0, 24.5, 32.4, 0, 26.07, 24.2)
DATA <- data.frame(MONTH, Group, lsmean, lower.CL, upper.CL)
DATA
```

Click on 'CODE' to see how the data is typed into R - use the same principle for your own data.

```{r FIG_R0, eval=FALSE}
```


\  

### The Basic

This is the most simple plot you can make. This is the essential part of the code for making the plot.

```{r FIG_R1, echo=FALSE, fig.align='center'}
library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group)) +
	geom_line(aes(linetype = Group)) +
	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL)) 

```

Show the code by clicking on 'CODE'.

```{r FIG_R1, eval=FALSE}
```


\  

### The Published

To replicate the published figure, some changes in the layout are needed. 


```{r FIG_R2, echo=FALSE, fig.align='center', fig.height=4.02, fig.width=5.36, dpi=300, fig.cap=""}

library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")

pd <- position_dodge(width = 0.1)

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group)) +
	geom_line(aes(linetype = Group), size=0.8, position = pd) +
	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.7, position = pd, linetype = 1) +
	geom_point(size = 2, position = pd, color = "black") +
	geom_hline(yintercept = 0) +
	scale_x_continuous(name="Time (months)", limits=c(-0.1,25), breaks=c(0,3,24)) +
	scale_y_continuous(name=expression("Change in KOOS"[5]), limits=c(-5, 40), 
				 breaks=c(-5,0,5,10,15,20,25,30,35,40)) +
	theme(panel.background = element_rect(fill = "white"),
		axis.line.x = element_line(colour = "black", size = 1),
		axis.line.y = element_line(colour = "black", size = 1), 
		legend.position="none")

```

To fit this figure to your own data, you may want to change the `name`, `limits` and `breaks` specifications in the 'scale_x_continuous' and 'scale_y_continuous' functions.

```{r FIG_R2, eval=FALSE}
```


\  

### The Wild Ones

Need something special for a powerpoint presentation? I have got you covered. The colors are easy to change.

\ 

```{r FIG_R3, echo=FALSE, fig.align='center'}

library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")

pd <- position_dodge(width = 0.1)

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group, colour = Group)) +
	geom_line(size=2, position = pd) +
	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 2, position = pd, linetype = 3) +
	geom_point(size = 2, position = pd) +
	geom_hline(yintercept = 0, size=1, colour = "darkgrey") +
	scale_x_continuous(name="Time (months)", limits=c(-0.1,25), breaks=c(0,3,24)) +
	scale_y_continuous(name=expression("Change in KOOS"[5]), limits=c(-5, 40), 
				 breaks=c(-5,0,5,10,15,20,25,30,35,40)) +
	scale_color_manual(values=c('green','deeppink1'))+
	theme(axis.title.x = element_text(colour = "white"),
		axis.title.y = element_text(colour = "white")) +
	theme(plot.background = element_rect(fill = "black"),
		panel.background = element_rect(fill = "black"),
		panel.grid.major = element_blank(), 
		panel.grid.minor = element_blank(),
		axis.line.x = element_line(colour = "darkgrey", size = 1),
		axis.line.y = element_line(colour = "darkgrey", size = 1), 
		legend.position="none") +
	annotate(geom="text", x=21, y=23, label="AMP", colour="green", size=5, fontface="bold", angle=10) +
   	annotate(geom="text", x=21, y=16, label="Placebo", colour="deeppink1", size=5, fontface="bold", angle=-2)

```

```{r FIG_R3, eval=FALSE}
```


\ 

```{r FIG_R4, echo=FALSE, fig.align='center'}
library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")
library(grid) #If the grid package is not installed, run this: install.packages("grid")

grad = colorRampPalette(c("red", "yellow"))(10)

pd <- position_dodge(width = 0.1)

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group)) +
  	annotation_custom(rasterGrob(grad, width=unit(1,"npc"), height=unit(1,"npc"))) +
	geom_line(aes(linetype = Group), size = 1, position = pd, colour = "white") +
  	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 2, position = pd, linetype = 3, color = "white") +
	geom_point(size = 1, position = pd, color = "white") +
  	geom_hline(yintercept = 0, size = 0.8, colour = "white") +
  	geom_text(aes(label = as.factor(lsmean))) +
	scale_x_continuous(name="Time (months)", limits=c(-0.1,25), breaks=c(0,3,24)) +
	scale_y_continuous(name=expression("Change in KOOS"[5]), limits=c(-5, 40), 
				 breaks=c(-5,0,5,10,15,20,25,30,35,40)) +
	theme(legend.position="none",
		axis.line.x = element_line(colour = "white", size = 2),
		axis.line.y = element_line(colour = "white", size = 2)) 

```

```{r FIG_R4, eval=FALSE}
```



