---
title: "R Graphs"
author: "by S. M. Nielsen"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}

library(knitr)    # For knitting document and include_graphics function
library(ggplot2)  # For plotting
library(png)      # For grabbing the dimensions of png files

```

\ 

# Line chart with confidence intervals - published in BMJ (BMJ style)

Let's try to reproduce the lower part in this figure published in BMJ by <a href="https://www.bmj.com/content/368/bmj.m697" target="_blank">Xu et al. (2020)</a>, where the efficacy  of manual acupuncture as prophylactic treatment for patients with episodic migraine was assessed. The graph shows the number of migraine days per 4 weeks (top), and the frequency of migraine attacks per 4 weeks (bottom) during the study period. Further details can be found in the article.

\  
\ 
<center>
```{r echo=FALSE}
img1_path <- "images//FIGURE_BMJ_2020_Migraine.png"
include_graphics(img1_path)
```
</center>

\ 

###Read in the data

To make it simple, all data needed for the figure is typed into R in this format.

```{r FIG_B0, echo=FALSE}
TIMElabel <- c("Baseline","1-4","5-8","9-12","13-16","17-20")
TIME <- c(1:6)
EST_manual <- c(3.8,2.7,2,1.8,1.7,1.4)
EST_sham <- c(4.1,3.1,2.4,2.3,2.3,2.5)
EST_usual <- c(3.7,3.5,3,3.3,3.0,3.3)
UL_manual <- c(5.1,4.0,3.3,3.3,3.0,2.7)
LL_manual <- c(2.4,1.3,0.7,0.3,0.3,0.1)
UL_sham <- c(6.7,4.7,3.8,3.8,4.0,4.1)
LL_sham <- c(1.5,1.4,1.0,0.8,0.6,1.0)
UL_usual <- c(5.0,4.9,4.2,4.6,4.0,4.7)
LL_usual <- c(2.4,2.1,1.8,2.0,2.0,1.8)

DATA <- as.data.frame(cbind(TIME, EST_manual,EST_sham,EST_usual,
                            UL_manual,UL_sham,UL_usual,
                            LL_manual,LL_sham,LL_usual))
DATA
```

Click on 'CODE' to see how the data is typed into R - use the same principle for your own data.
(Please note, as I was not involved in this project, the values are simply just read from the published figure)

```{r FIG_B0, eval=FALSE}
```


\  

### The Basic

This is the most simple plot you can make. This is the essential part of the code for making the plot.

```{r FIG_X1, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
require(plotrix) #If the plotrix package is not installed, run this: install.packages("plotrix")

#Plot
plot(DATA$TIME, DATA$EST_manual,
     ylim=c(0,8), xlim=c(0,6), xaxt='n',
     ylab="No of migraine attacks per 4 weeks", xlab="Time (week)")

#X axis
axis(1, c(1:6), labels=TIMElabel)

#Lines and CIs
lines(DATA$TIME, DATA$EST_usual, col="red")
plotCI(DATA$TIME, DATA$EST_usual, ui=UL_usual, li=LL_usual, add=TRUE, col="red")

lines(DATA$TIME, DATA$EST_sham, col="orange")
plotCI(DATA$TIME, DATA$EST_sham, ui=UL_sham, li=LL_sham, add=TRUE, col="orange")

lines(DATA$TIME, DATA$EST_manual, col="purple")
plotCI(DATA$TIME, DATA$EST_manual, ui=UL_manual, li=LL_manual, add=TRUE, col="purple")

legend("top", legend=c("Manual acupuncture", "Sham acupuncture", "Usual care"),
       col=c("purple", "orange", "red"), lwd=1)

```

Show the code by clicking on 'CODE'.

```{r FIG_X1, eval=FALSE}
```


\  

### The Published

To replicate the published figure, some changes in the layout are needed. 


```{r FIG_X2, echo=FALSE, fig.align='center', fig.cap="", fig.height=5, fig.width=7, message=FALSE, warning=FALSE, dpi=400}

library(plotrix) #If the plotrix package is not installed, run this: install.packages("plotrix")
library(berryFunctions) #If the berryFunctions package is not installed, run this: install.packages("berryFunctions")

#Axis min and max
YMAX <- 8; YMIN <- 0
XMAX <- 6.5; XMIN <- 0.5

#Plot
plot(DATA$TIME, DATA$EST_manual, type="o", pch=20,
     ylim=c(YMIN,YMAX+1), xlim=c(XMIN,XMAX), yaxs="i", xaxs="i", axes=F, 
     ylab="", xlab="")

#Axes and labels
axis(2, c(0,2,4,6,8), col = NA, col.ticks = NA, las=2, line=-0.8)
axis(1, c(1:6), labels=TIMElabel, col = NA, col.ticks = "grey", tck=-0.04, line=-0.7)
mtext("No of migraine attacks per 4 weeks", 2, 2, font=2)
mtext("Time (week)", 1, 1.5, font=2, at=XMAX, adj=1)

#Plotting area
roundedRect(XMIN,YMIN,XMAX,YMAX, rounding=0.05, border="grey") 
abline(h=c(2,4,6), col="grey")

#Lines and CIs
lines(DATA$TIME, DATA$EST_usual,  type="o", pch=20, lwd=3, cex=1.7, lty=3, col="deeppink3")
plotCI(DATA$TIME, DATA$EST_usual, ui=UL_usual, li=LL_usual, add=TRUE, lwd=3, sfrac=0.006, col="deeppink3")

lines(DATA$TIME, DATA$EST_sham,   type="o", pch=20, lwd=3, cex=1.7, lty=2, col="darkgoldenrod1")
plotCI(DATA$TIME, DATA$EST_sham, ui=UL_sham, li=LL_sham, add=TRUE, lwd=3, sfrac=0.006, col="darkgoldenrod1")

lines(DATA$TIME, DATA$EST_manual, type="o", pch=20, lwd=3, cex=1.7, col="mediumpurple3")
plotCI(DATA$TIME, DATA$EST_manual, ui=UL_manual, li=LL_manual, add=TRUE, lwd=3, sfrac=0.006, col="mediumpurple3")


legend(XMIN, 9, legend=c("Manual acupuncture", "Sham acupuncture", "Usual care"),
       col=c("mediumpurple3", "darkgoldenrod1", "deeppink3"), 
       lty=c(1,2,3), 
       cex=0.8, lwd=3, pch=20,
       box.lty=0, horiz=TRUE)

```

To fit this figure to your own data, you may want to change the values saved for `XMIN`,`XMAX`, `YMIN` and `YMAX`, the position of the ticks on the x-axis which currently reads `c(0,2,4,6,8)`,  the position of the horizontal lines plotting area which reads `h=c(2,4,6)`, and the position of the legend which currently reads `8.7` on the y axis.

```{r FIG_X2, eval=FALSE}
```


\  


\ 
\ 



# Line chart with confidence intervals - published in BMJ Open

Let's try to reproduce panel A in this figure published in BMJ Open by <a href="https://www.ncbi.nlm.nih.gov/pubmed/29420232" target="_blank">Roos et al. (2018)</a>, where knee surgery (arthroscopic partial meniscectomy) was compare to sham surgery (skin incision only). The graph shows the mean change from baseline at 3 and 24 months in the primary outcome (KOOS~5~ score, a higher score is better). Further details can be found in the article.

\  
\ 
<center>
```{r echo=FALSE}
img1_path <- "images//FIGURE_Roos_2018.png"
include_graphics(img1_path)
```
</center>

\ 

###Read in the data

To make it simple, all data needed for the figure is typed into R in this format.

```{r FIG_R0, echo=FALSE}
MONTH <- c(0,3,24,0,3,24)
Group <- c("APM","APM","APM","Placebo","Placebo","Placebo")
lsmean <- c(0, 13.9, 21.8, 0, 15.1, 13.6)
lower.CL <- c(0, 3.2, 11.2, 0, 4.2, 3.0)
upper.CL <- c(0, 24.5, 32.4, 0, 26.07, 24.2)
DATA <- data.frame(MONTH, Group, lsmean, lower.CL, upper.CL)
DATA
```

Click on 'CODE' to see how the data is typed into R - use the same principle for your own data.

```{r FIG_R0, eval=FALSE}
```


\  

### The Basic

This is the most simple plot you can make. This is the essential part of the code for making the plot.

```{r FIG_R1, echo=FALSE, fig.align='center'}
library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group)) +
	geom_line(aes(linetype = Group)) +
	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL)) 

```

Show the code by clicking on 'CODE'.

```{r FIG_R1, eval=FALSE}
```


\  

### The Published

To replicate the published figure, some changes in the layout are needed. 


```{r FIG_R2, echo=FALSE, fig.align='center', fig.height=4.02, fig.width=5.36, dpi=300, fig.cap=""}

library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")

pd <- position_dodge(width = 0.1)

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group)) +
	geom_line(aes(linetype = Group), size=0.8, position = pd) +
	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.7, position = pd, linetype = 1) +
	geom_point(size = 2, position = pd, color = "black") +
	geom_hline(yintercept = 0) +
	scale_x_continuous(name="Time (months)", limits=c(-0.1,25), breaks=c(0,3,24)) +
	scale_y_continuous(name=expression("Change in KOOS"[5]), limits=c(-5, 40), 
				 breaks=c(-5,0,5,10,15,20,25,30,35,40)) +
	theme(panel.background = element_rect(fill = "white"),
		axis.line.x = element_line(colour = "black", size = 1),
		axis.line.y = element_line(colour = "black", size = 1), 
		legend.position="none")

```

To fit this figure to your own data, you may want to change the `name`, `limits` and `breaks` specifications in the 'scale_x_continuous' and 'scale_y_continuous' functions.

```{r FIG_R2, eval=FALSE}
```


\  

### The Wild Ones

Need something special for a powerpoint presentation? I have got you covered. The colors are easy to change.

\ 

```{r FIG_R3, echo=FALSE, fig.align='center'}

library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")

pd <- position_dodge(width = 0.1)

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group, colour = Group)) +
	geom_line(size=2, position = pd) +
	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 2, position = pd, linetype = 3) +
	geom_point(size = 2, position = pd) +
	geom_hline(yintercept = 0, size=1, colour = "darkgrey") +
	scale_x_continuous(name="Time (months)", limits=c(-0.1,25), breaks=c(0,3,24)) +
	scale_y_continuous(name=expression("Change in KOOS"[5]), limits=c(-5, 40), 
				 breaks=c(-5,0,5,10,15,20,25,30,35,40)) +
	scale_color_manual(values=c('green','deeppink1'))+
	theme(axis.title.x = element_text(colour = "white"),
		axis.title.y = element_text(colour = "white")) +
	theme(plot.background = element_rect(fill = "black"),
		panel.background = element_rect(fill = "black"),
		panel.grid.major = element_blank(), 
		panel.grid.minor = element_blank(),
		axis.line.x = element_line(colour = "darkgrey", size = 1),
		axis.line.y = element_line(colour = "darkgrey", size = 1), 
		legend.position="none") +
	annotate(geom="text", x=21, y=23, label="AMP", colour="green", size=5, fontface="bold", angle=10) +
   	annotate(geom="text", x=21, y=16, label="Placebo", colour="deeppink1", size=5, fontface="bold", angle=-2)

```

```{r FIG_R3, eval=FALSE}
```


\ 

```{r FIG_R4, echo=FALSE, fig.align='center'}
library(ggplot2) #If the ggplot2 package is not installed, run this: install.packages("ggplot2")
library(grid) #If the grid package is not installed, run this: install.packages("grid")

grad = colorRampPalette(c("red", "yellow"))(10)

pd <- position_dodge(width = 0.1)

ggplot(DATA, aes(x = MONTH, y = lsmean, group = Group)) +
  	annotation_custom(rasterGrob(grad, width=unit(1,"npc"), height=unit(1,"npc"))) +
	geom_line(aes(linetype = Group), size = 1, position = pd, colour = "white") +
  	geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 2, position = pd, linetype = 3, color = "white") +
	geom_point(size = 1, position = pd, color = "white") +
  	geom_hline(yintercept = 0, size = 0.8, colour = "white") +
  	geom_text(aes(label = as.factor(lsmean))) +
	scale_x_continuous(name="Time (months)", limits=c(-0.1,25), breaks=c(0,3,24)) +
	scale_y_continuous(name=expression("Change in KOOS"[5]), limits=c(-5, 40), 
				 breaks=c(-5,0,5,10,15,20,25,30,35,40)) +
	theme(legend.position="none",
		axis.line.x = element_line(colour = "white", size = 2),
		axis.line.y = element_line(colour = "white", size = 2)) 

```

```{r FIG_R4, eval=FALSE}
```



