---
title: "R Meta"
author: "by S. M. Nielsen"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}

library(knitr)    # For knitting document and include_graphics function
library(metafor)  # For meta-analysis and plotting
library(png)      # For grabbing the dimensions of png files

```

\ 

# Meta-analysis and forest plot - published in Cochrane review (Cochrane style)

Let's try to reproduce first meta-analysis and forest plot presented in the Cochrane review by <a href="https://www.cochranelibrary.com/cdsr/doi/10.1002/14651858.CD013273/full" target="_blank">Leopoldino et al. (2019)</a>, where the benefits and harms of paracetamol versus placebo for knee and hip osteoarthritis was investigated.

The forest plot shows the effect on pain (expressed on scale of 0 to 100) at immediate term (follow-up two weeks or less) or short term (follow-up more than two weeks but three months or less). Further details can be found in the article.

\  
\ 
<center>
```{r echo=FALSE}
img1_path <- "images//FIGURE_Cochrane_Leopoldino_2019.png"
include_graphics(img1_path)
```
</center>

\ 

###Read in the data

To make it simple, all data needed for the figure is typed into R in this format. 
(Data can often be downloaded directly from Cochrane reviews with the option 'Download statistical data' and read using 'Review Manager', from which it can be copy-pasted to e.g. Excel)

```{r FIG_B0, echo=FALSE}
Study <- c("Case 2003","Golden 2004","Miceli-Richard 2004","Prior 2014","Zoppi 1995","Altman 2007 (1)","Altman 2007 (2)","Case 2003","Herrero-Beaumont 2007","Miceli-Richard 2004","Pincus 2004a","Pincus 2004b","Prior 2014")

Term <- c("Immediate","Immediate","Immediate","Immediate","Immediate","Short","Short","Short","Short","Short","Short","Short","Short")

Mean_I <- as.numeric(c(-0.9,-22.2,-16,-26.4,-20,-22.8,-26.5,-4.8,-2.4,-23.6,-17.4,-13.8,-30))
SD_I <- c(11.7,18.1,21,24.2,21.5,21.6,25.5,16.6,3.1454,21.7,26,23.7,20.9)
N_I <- c(27,145,385,267,28,158,160,22,108,298,171,185,177)
Mean_C <- c(-0.3,-18,-15,-20.5,-10.8,-19.6,-19.6,-3.1,-1.8,-22.8,-10.5,-7.6,-25.8)
SD_C <- c(10.5,16.5,21,24.5,18,22.5,22.5,19.7,3.5994,21.7,25.2,26.9,20.3)
N_C <- c(26,149,356,275,28,82,83,19,104,262,172,182,172)

DATA <- as.data.frame(cbind(Study, Term, Mean_I, SD_I, N_I, Mean_C, SD_C, N_C))

#Change to correct data type
DATA[,c(3:8)] <- sapply(DATA[,c(3:8)],as.character)
DATA[,c(3:8)] <- sapply(DATA[,c(3:8)],as.numeric)

DATA
```

Click on 'CODE' to see how the data is typed into R - use the same principle for your own data.

```{r FIG_B0, eval=FALSE}
```



\  

### The Basic

This is the most simple plot you can make. This is the essential part of the code for making the plot.

```{r FIG_X1, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
require(metafor) #If the metafor package is not installed, run this: install.packages("metafor")

MA_DATA <- escalc(measure="MD", m1i=Mean_I, sd1i=SD_I, m2i=Mean_C, sd2i=SD_C, n1i=N_I, n2i=N_C, data=DATA)

rma(yi, vi, data=subset(MA_DATA, Term=="Immediate"), method="REML")
rma(yi, vi, data=subset(MA_DATA, Term=="Short"), method="REML")	#WHY NOT EXACT SAME??

http://www.metafor-project.org/doku.php/plots:forest_plot_with_subgroups
https://www.rdocumentation.org/packages/metafor/versions/2.4-0/topics/forest.rma

# subsetting??

```

Show the code by clicking on 'CODE'.

```{r FIG_X1, eval=FALSE}
```


\  